---
title: "R Notebook"
output: html_notebook
---

# Load libraries
```{r, message=F, warning=F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(nlme)
options(scipen=999)
```

# Load daily data
```{r, message=F, warning=F, results="hold"}
dailyDat <- read.table("/Users/Santi/Desktop/rProjects/dissertation/analysis/data.tsv",
                  header = T,
                  sep = "\t",
                  stringsAsFactors = F,
                  na.strings = c(999994, 999995, 999996, 999998, 98.00))
```

# Number of Participants
```{r}
nParticipants <- dailyDat %>%
  dplyr::select(M2ID) %>%
  n_distinct()
```


# Select cortisol variables of interest from daily data
```{r, message=F, warning=F, results="hold"}
cortTimesDat <- dailyDat %>%
  dplyr::select(id = M2ID, gender = B1PGENDER, age = B1PAGE_M2, medUseNone1Yes = B2DMED10, startWork = B2DA3AAP, endWork = B2DA3BAP, providedCort1yes2no = B2DCORT, day = B2DDAY, wakeCort = B2DCORWT, wakeThirtyCort = B2DCORAT, noonCort = B2DCORLT, pmCort = B2DCORBT) %>%
  dplyr::filter(providedCort1yes2no == 1) %>%
  dplyr::filter(day == 2 | day == 3 | day == 4 | day == 5) %>%
  pivot_longer(9:12, names_to = "cortTimeText", values_to = "cortTime") 
  
cortValuesDat <- dailyDat %>%
  dplyr::select(id = M2ID, providedCort1yes2no = B2DCORT, day = B2DDAY, wakeCort = B2DCORW, wakeThirtyCort = B2DCORA, noonCort = B2DCORL, pmCort = B2DCORB) %>%
  dplyr::filter(providedCort1yes2no == 1) %>%
  dplyr::filter(day == 2 | day == 3 | day == 4 | day == 5) %>%
  pivot_longer(4:7, names_to = "cortTimeText", values_to = "cortNmol") %>%
  dplyr::select(id, day, cortTimeText, cortNmol)

cortDat <- left_join(cortTimesDat, cortValuesDat, by = c("id", "day", "cortTimeText"))

str(cortDat)
```

# Remove NAs from cortTime and cortNmol
```{r, message=F, warning=F, results="hold"}
naCortNmolPercent <- round((sum(is.na(cortDat$cortNmol)) / nrow(cortDat)) * 100, 2)
naCortTimePercent <- round((sum(is.na(cortDat$cortTime)) / nrow(cortDat)) * 100, 2)

naCortNmolSum <- sum(is.na(cortDat$cortNmol))
naCortTimeSum <- sum(is.na(cortDat$cortTime))

masterCortDat <- cortDat %>%
  dplyr::filter(!is.na(cortNmol), !is.na(cortTime))

str(masterCortDat)
```

# Remove samples collected outside of 24hr window
```{r, message=F, warning=F, results="hold"}
samplesOutside24hrWindow <- masterCortDat %>%
  dplyr::filter(cortTime < 0.00 | cortTime > 24.00) %>%
  nrow()
#607 samples collected prior to beginning of of interview day or after end of interview day

samplesOutside24hrPercent <- round((samplesOutside24hrWindow / nrow(masterCortDat)) * 100, 2)


masterCortDat <- masterCortDat %>%
  dplyr::filter(cortTime >= 0.00, cortTime <= 24.00)

str(masterCortDat)
```

# Remove samples with cortNmol > 60
```{r, message=F, warning=F, results="hold"}
masterCortDat %>%
  dplyr::filter(cortNmol > 60)
#381 greater than 60

greaterThanSixtyPercent <- round((nrow(subset(masterCortDat, cortNmol > 60)) / nrow(masterCortDat)) * 100, 2)

greaterThanSixtySum <- nrow(subset(masterCortDat, cortNmol > 60))

masterCortDat <- masterCortDat %>%
  dplyr::filter(cortNmol <= 60)

str(masterCortDat)
```

# Remove waking cortisol samples collected before 0400h or after 1200h
```{r, message=F, warning=F, results="hold"}
earlyLateWakeCortSum <- masterCortDat %>%
  dplyr::filter(cortTimeText == "wakeCort") %>%
  dplyr::filter(cortTime < 4.00 | cortTime > 12) %>%
  nrow()
#173 wake cort collected before 4am or after 12pm

earlyLateWakeCortPercent <- round((earlyLateWakeCortSum / nrow(masterCortDat)) * 100, 2)

# df that contains wakeCortTime between 4 and 12
wakeCortKeepMe <- masterCortDat %>%
  dplyr::filter(cortTimeText == "wakeCort") %>%
  dplyr::filter(between(cortTime, 4.00, 12.00)) 

# remove all wakeCort rows from master dat
masterCortDat <- masterCortDat %>%
  dplyr::filter(cortTimeText != "wakeCort")

# bind wakeCortKeepMe rows w/ masterCortDat rows
masterCortDat <- bind_rows(masterCortDat, wakeCortKeepMe)

# arrange rows
masterCortDat <- masterCortDat %>%
  arrange(id, day, cortTime)

str(masterCortDat)
```

# Number of 1200h Values that are 10 nmol/L greater than the wake + 30 values?
```{r, message=F, warning=F, results="hold"}
erronLunchValues <- masterCortDat %>% 
  mutate(dummy1 = case_when(
    cortTimeText == "wakeThirtyCort" ~ 1,
    cortTimeText == "noonCort" ~ 1,
    TRUE ~ 0
  )) %>%
  dplyr::filter(dummy1 == 1) %>%
  arrange(id, day, cortTime) %>%
  group_by(id, day) %>%
  mutate(dummyDif = last(cortNmol) - first(cortNmol)) %>%
  mutate(assignDummy = case_when(
    dummyDif > 10 ~ 1,
    TRUE ~ 0
  )) %>%
  dplyr::filter(assignDummy == 1) %>%
  nrow()
# 166 samples

erronLunchPercent <- round((erronLunchValues / nrow(masterCortDat)) * 100, 2)

nmolWakeNoon10 <- masterCortDat %>%
  mutate(dummy1 = case_when(
    cortTimeText == "wakeThirtyCort" ~ 1,
    cortTimeText == "noonCort" ~ 1,
    TRUE ~ 0
  )) %>%
  dplyr::filter(dummy1 == 1) %>%
  arrange(id, day, cortTime) %>%
  group_by(id, day) %>%
  mutate(dummyDif = last(cortNmol) - first(cortNmol)) %>%
  mutate(assignDummy = case_when(
    dummyDif > 10 ~ 1,
    TRUE ~ 0
  )) %>%
  dplyr::filter(assignDummy == 1) %>%
  dplyr::select(-dummy1, -dummyDif, -assignDummy)

masterCortDat <- anti_join(masterCortDat, nmolWakeNoon10)

str(masterCortDat)
```

# How many participants are using medications that impact HPA axis function,
#didn't know or did not answer this question?
```{r, message=F, warning=F, results="hold"}
masterCortDat %>%
  dplyr::filter(medUseNone1Yes == 2 | medUseNone1Yes == 8 | medUseNone1Yes == 9) %>%
  dplyr::select(id) %>%
  distinct() %>%
  nrow()
# 895 of 1731

masterCortDat %>%
  dplyr::select(id) %>%
  n_distinct()
```

# Compute natural logarithm of cortisol
```{r, message=F, warning=F, results="hold"}
masterCortDat <- masterCortDat %>%
  mutate(logCort = log(cortNmol))

str(masterCortDat)
```

# Select emotion data and join with hypotheses 1 & 2 cortisol data
```{r, message=F, warning=F, results="hold"}
emoDat <- dailyDat %>%
  dplyr::select(id = M2ID, day = B2DDAY, numberOfStressors = B2DN_STR, stressorNegAffSum = B2DSNEGS, stressorNegAffMean = B2DSNEGA, stressorSeveritySum = B2DSSEVS, stressorSeverityMean = B2DSSEVA,  dailyMeanPosAff = B2DPOSAV, dailyMeanNegAff = B2DNEGAV, stressorControlSum = B2DSCOTS, stressorControlMean = B2DSCOTA, numberOfPhysicalSx = B2DN_SYM, physicalSxSeverity = B2DSYMAV, dailyInGoodSpirits = B2DC7, dailyCheerful = B2DC8, dailyExtremelyHappy = B2DC9, dailyCalmPeaceful = B2DC10, dailySatisfied = B2DC11, dailyFullOfLife = B2DC12, dailyCloseToOthers = B2DC21, dailyLikeYouBelong = B2DC22, dailyEnthusiastic = B2DC23, dailyAttentive = B2DC24, dailyProud = B2DC25, dailyActive = B2DC26, dailyConfident = B2DC27, dailyRestlessFidgety = B2DC1, dailyNervous = B2DC2, dailyWorthless = B2DC3, dailySoSadNoCheerUp = B2DC4, dailyEverythingAnEffort = B2DC5, dailyHopeless = B2DC6, dailyLonely = B2DC13, dailyAfraid = B2DC14, dailyJittery = B2DC15, dailyIrritable = B2DC16, dailyAshamed = B2DC17, dailyUpset = B2DC18, dailyAngry = B2DC19, dailyFrustrated = B2DC20) %>%
  dplyr::filter(day == 2 | day == 3 | day == 4 | day == 5)
```

# Select variables of interest from midus 2 data
```{r, message=F, warning=F, results="hold"}
midusDat <- read.table("/Users/Santi/Desktop/rProjects/dissertation/analysis/midusTwoData.tsv",
                  header = T,
                  sep = "\t",
                  stringsAsFactors = F)

midusDat <- midusDat %>%
  dplyr::select(id = M2ID, ethnicity = B1PF7A, education = B1PB1, smoker = B1PA39, somaticAmplification = B1SAMPLI, neuroticism = B1SNEURO, extraversion = B1SEXTRA, opennessToExp = B1SOPEN, conscientiousness = B1SCONS1, agreeableness = B1SAGREE, agency = B1SAGENC, mentalHealthTx = B1SUSEMH, mastery = B1SMASTE, constraints = B1SCONST,
                control = B1SCTRL, growth = B1SREINT, activeCoping = B1SACTIV,
                planning = B1SPLAN, venting = B1SVENT, denial = B1SDENIA,
                disengage = B1SDISEN, foodCoping = B1SFDCOP, problemFocused = 
                  B1SPRCOP, emotionFocused = B1SEMCOP, seekSupport = B1SSUFFI,
                friendSupport = B1SFDSPO, friendStrain = B1SFDSNE, compositeFriend =
                  B1SFDSOL, familySupport = B1SKINPO, familyStrain = B1SKINNE,
                compositeFamily = B1SFAMSO)
```

# Join dataframes
```{r, message=F, warning=F, results="hold"}
masterCortDat <- left_join(masterCortDat, midusDat, by = "id") %>%
  dplyr::select(id, age, gender, ethnicity, education, smoker, medUseNone1Yes,
         opennessToExp, conscientiousness, agreeableness, extraversion,
         neuroticism, somaticAmplification, agency, mentalHealthTx, mastery,
         constraints, control, growth, activeCoping, planning, venting, denial,
         disengage, foodCoping, problemFocused, emotionFocused, seekSupport, 
         friendSupport, friendStrain, compositeFriend, familySupport, familyStrain,
         compositeFamily, day, cortTimeText, cortTime, cortNmol, logCort)

str(masterCortDat)

masterCortDat <- left_join(masterCortDat, emoDat, by = c("id", "day"))

numberOfCortSamples <- nrow(masterCortDat)

write.csv(masterCortDat, "big5Dat.csv", row.names = F)
```






































