---
output:
  html_document: default
  pdf_document: default
  word_document: default
bibliography: refs.bib
nocite: | 
  @larsen1991day
---

```{r, code = readLines(knitr::purl("/Users/Santi/Desktop/rProjects/dissertation/analysis/big5DataWrangling.Rmd")), eval=TRUE, include=F}
```

```{r, code = readLines(knitr::purl("/Users/Santi/Desktop/rProjects/dissertation/analysis/big5DataInspection.Rmd")), eval=TRUE, include=F}
```

# **CHAPTER 2**  

# **METHOD**

## **Sample and Procedure**

The present study used data from the National Study of Daily Experiences (NSDE), a substudy of the second wave of the Midlife in the United States Survey (MIDUS II). A nonclinical sample of participants ($N = `r nParticipants`$) completed 10-15 minute semi-structured evening telephone interviews for eight consecutive days, where daily physical symptoms were assessed alongside other variables of interest (see Almeida et al. 2002 for methodological details). However, saliva samples to be assayed for cortisol were only collected on days two through five, four times per day. Informed consent was obtained from all participants. Participants were compensated $20 for participating in the NSDE. Per the cortisol exclusion criteria outlined below, data from `r analyticSample` participants were analyzed in the present study. 

Age of participation ranged from `r ageMin` to `r ageMax` years ($M = `r ageMean`, SD = `r ageSd`$) and `r femalePercent`% of participants identified as female. As shown in table 1, `r ethnFreqCaucPercent`% of participants identified as Caucasian, `r ethnFreqAfAmPercent`% identified as African American, `r ethnFreqAsianPercent` % identified as Asian, `r ethnFreqNativePercent`% identified as Native American/Alaska Native/Aleutian Islander/Eskimo and `r ethnFreqOtherPercent`% identified as other. As shown in Table 1, `r eduFreqHsGradGedPercent`% had earned a high school diploma or a GED, `r manuscriptCollegePercent`% had at least some college education or a college degree and `r manuscriptGradPercent` had at least some graduate education or a graduate degree.

## **Measures**

### **Personality**

Personality was assessed with a self-report measure specifically developed for the MIDUS studies [@lachman1997midlife]. Six factors were derived from 30 items: agency, agreeableness, neuroticism, openness to experience, conscientiousness and extraversion [@lachman1997midlife]. The present analysis excluded agency in that it is not a factor in the Five Factor theory of personality. Participants were asked to rate how much each of the 30 adjectives described them on a 4-point Likert-type scale ranging from "not at all" and "a little" to "some" and "a lot" [@lachman1997midlife]. The mean of each factor was computed, with higher scores reflecting greater endorsement of the trait. The agreeableness factor included, "helpful", "warm", "caring", "softhearted" and "sympathetic" (Cronbach’s $\alpha$ = .80); the neuroticism factor included, "moody", "worrying", "nervous" and "not calm" (Cronbach’s $\alpha$ = .74); the openness to experience factor included, "creative", "imaginative", "intelligent", "curious", "broadminded", "sophisticated" and "adventurous" (Cronbach’s $\alpha$ = .77); the conscientiousness factor included, "organized", "responsible", "hardworking" and "careless" (Cronbach’s $\alpha$ = .58); the extraversion factor included, "outgoing", "friendly", "lively", "active" and "talkative" (Cronbach’s $\alpha$ = .78) [@lachman1997midlife]. 

### **Number of Physical Symptoms and Physical Symptom Severity**
Number of physical symptoms was assessed each day with an adapted version of the Larsen and Kasimatis (1991) symptom checklist. The inventory consists of 28 physical symptoms: “Headache”, “Backache”, “Muscle Soreness”, “Fatigue”, “Joint Pain”, “Muscle Weakness”, “Cough”, “Sore Throat”, “Fever”, “Chills”, “Other Cold Or Flu Symptoms”, “Nausea”, “Allergies”, “Diarrhea”, “Constipation”, “Poor Appetite”, “Other Stomach Problems”, “Chest Pain”, “Dizziness”, “Shortness of Breath”, “Menstrual Related Symptoms”, “Hot Flashes or Flushes”, “Any Other Physical Symptoms”, “Skin Related Symptoms”, “Eye Related Symptoms”, “Ear Related Symptoms”, “Teeth Related Symptom”, “Leg or Foot Related Symptom”. Participants rated each item using a binary Yes/No scale. After recoding items so that 0 represents "No" and 1 represents "Yes", items were summed for a measure of total number of daily symptoms. To measure physical symptom severity, participants rated the severity of each physical symptom endorsed in the above checklist on a 10-point Likert-type scale, with 1 indicating "Very Mild" and 10 indicating "Very Severe". The average of all severity ratings was used as a measure of daily physical symptom severity. In the present study, we used the average of days two through five of both measures in all analyses. 

### **Salivary Cortisol**

Using at-home cotton swab Salivette collection devices (Sarstedt, Nümbrecht, Germany), participants provided four saliva samples per day for four consecutive days [@Almeida2002-fo]. Participants were instructed to provide samples at wake, 30-minutes post-wake, before lunch and before bed [@Almeida2002-fo]. Participants were also instructed to refrain from consuming a meal at least one hour prior to collection, drinking or eating dairy products 20 minutes prior to collection, brushing their teeth and consuming caffeine 30 minutes prior to collection (alcohol?). In addition, participants recorded all prescription and over-the-counter medications and disclosed any history of endocrine-related disorders.

Intra- and inter-coefficients of variation less than 5% were derived from the luminescence immunoassay procedure [@Almeida2002-fo]. After collecting a total of 16 samples, participants shipped the Salivette devices to the MIDUS Biological Core at the University of Wisconsin. At MIDUS Biological Core, samples were stored at -60C before being thawed, centrifuged and assayed for cortisol [@Almeida2002-fo]. Also, each cortisol sample was examined and corrected if not within a 4 to 9 pH range.

In addition to paper-and-pencil logs and evening telephone interviews to record timestamps, ~25% of participants used an electronic box that recorded the time at which participants opened and closed the box to use a Salivette collection device [@Almeida2002-fo]. At each collection occasion, self-reported timestamps were correlated above .90. Correlations between self-reported time-stamps and the electronic box were correlated above .75, with a stronger correlation in the morning than in the evening [@Almeida2002-fo].

Based on prior studies on the NSDE dataset, the following exclusion criteria were used [@almeida2009interindividual; @charles2013wear; @piazza2013age; @Sin:2017gt]. First, cortisol samples with mising values were excluded ($n = `r naCortNmolSum`, `r naCortNmolPercent`$%). Second, cortisol values with missing timestamps ($n = `r naCortTimeSum`, `r naCortTimePercent`$%) or with concentration levels greater than 60 nmol/L were excluded ($n = `r greaterThanSixtySum`, `r greaterThanSixtyPercent`$%). Third, waking cortisol samples with timestamps before 0400h or after 1200h, indicating aberrant sleep patterns, were removed ($n = `r earlyLateWakeCortSum`, `r earlyLateWakeCortPercent`$%). Fourth, cortisol samples collected outside of the 24h window were removed ($n = `r samplesOutside24hrWindow`, `r samplesOutside24hrPercent`$%). Fifth, to avoid modeling erroneously high cortisol levels due to lunchtime meals, noon cortisol samples 10 nmol/L or greater than the wake +30 sample were removed ($n = `r erronLunchValues`, `r erronLunchPercent`$%). The final analytic sample included `r analyticSample` participants, who provided  `r numberOfCortSamples` saliva samples. (covariate missing data)

## **Statistical Analyses**

The R statistical computing language was used for all analyses. The natural logarithm of cortisol was computed to correct for skewness in the distribution. To avoid modeling the cortisol awakening response, the second salivary cortisol sample from day 2 was removed for each participant (i.e., wake + 30 minutes) [@Adam:2009et; @kraemer2006design]. The cortisol IIV variable was derived by developing a 2-level variance components model, with log-transformed cortisol entered as the outcome variable and time-since-waking entered as the predictor. Random intercept and slope variance components were estimated at level 2, while within-participant residual variability was estimated at level 1. Cortisol IIV was calculated as the SD of raw residuals (observed - fitted) at level 1. As a result, each participant had one, time-invariant, cIIV value. An alpha level of .05 was used for all model estimates. 

Covariates known to be associated with measures of salivary cortisol, including age, gender, steroid medication use, education level and smoking status, were individually tested in a robust regression model. Covariates that were not  associated with cIIV, which included gender and medication use, were not included in the models below. However, age, education and smoking status were associated with cIIV and were included as covariates in all models.

Due to collinearity and OLS assumption violations, seven robust regression models were estimated to test the hypotheses of the study. Cortisol IIV was modeled as a function of each personality factor and symptom severity variables, separately, while adjusting for the effects of age, education and smoking status. Huber M-estimation was chosen as the estimation method because it weighs deviations as the absolute value of the deviation, which renders the model less sensitive to outliers. All predictors were grand-mean-centered to aid interpretability of regression coefficients. The Benjamini-Hochberg procedure was used to correct for multiple comparisons [@benjamini1995controlling]. 

**Table 1.** Demographic Characteristics

| Variable       | Level                    | Frequency ($N = 1731$)               | %                              |
|----------------|--------------------------|-------------------------|--------------------------------|
| Gender         | Male                     | `r maleFreq`            | `r malePercent`                |
|                | Female                   | `r femaleFreq`          | `r femalePercent`              |
| Age            | 30-39 years old          | `r age3039Freq`         | `r age3039FreqPercent`         |
|                | 40-49 years old          | `r age4049Freq`         | `r age4049FreqPercent`         |
|                | 50-59 years old          | `r age5059Freq`         | `r age5059FreqPercent`         |
|                | 60-69 years old          | `r age6069Freq`         | `r age6069FreqPercent`         |
|                | 70-79 years old          | `r age7079Freq`         | `r age7079FreqPercent`         |
|                | 80 years and older       | `r age8089Freq`         | `r age8089FreqPercent`         |
| Education      | Elementary/Middle School | `r eduFreqElemMidSchl`  | `r eduFreqElemMidSchlPercent`  |
|                | Some High School         | `r eduFreqSomeHs`       | `r eduFreqSomeHsPercent`       |
|                | High School Graduate/GED | `r eduFreqHsGradGed`    | `r eduFreqHsGradGedPercent`    |
|                | Some College             | `r eduFreqSomeCollege`  | `r eduFreqSomeCollegePercent`  |
|                | Community College Degree | `r eduFreqComColDeg`    | `r eduFreqComColDegPercent`    |
|                | Bachelor's Degree        | `r eduFreqBachDeg`      | `r eduFreqBachDegPercent`      |
|                | Some Graduate School     | `r eduFreqSomeGradSchl` | `r eduFreqSomeGradSchlPercent` |
|                | Master's Degree          | `r eduFreqMasterDeg`    | `r eduFreqMasterDegPercent`    |
|                | Doctorate Degree         | `r eduFreqDocDeg`       | `r eduFreqDocDegPercent`       |
|                | Not available            | `r eduFreqNa`           | `r eduFreqNaPercent`           |
| Ethnicity      | Caucasian                | `r ethnFreqCauc`        | `r ethnFreqCaucPercent`        |
|                | African American         | `r ethnFreqAfAm`        | `r ethnFreqAfAmPercent`        |
|                | Native American          | `r ethnFreqNative`      | `r ethnFreqNativePercent`      |
|                | Asian                    | `r ethnFreqAsian`       | `r ethnFreqAsianPercent`       |
|                | Other                    | `r ethnFreqOther`       | `r ethnFreqOtherPercent`       |
|                | Not available            | `r ethnicFreqNa`        | `r ethnicFreqNaPercent`        |
| Medication Use | Yes                      | `r medNoFreq`           | `r medNoFreqPercent`           |
|                | No                       | `r medYesFreq`          | `r medYesFreqPercent`          |
|                | Not available            | `r medNa`               | `r medNaFreqPercent`           |
| Smoker         | Yes                      | `r smokerYesFreq`       | `r smokerYesFreqPercent`       |
|                | No                       | `r smokerNoFreq`        | `r smokerNoFreqPercent`        |
|                | Not available            | `r smokerFreqNa`        | `r smokerFreqNaPercent`        |
















