---
title: "Data Inspection"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Load libraries
```{r, message=F, warning=F}
library(tidyverse)
options(scipen=999)
```

# Load data
```{r, message=F, warning=F}
hyp1Dat <- read_csv("hyp1dat.csv")
hyp2Dat <- read_csv("hyp2Dat.csv")
```

# Assign NA for missing, do not know, inapplicable: panas, daily vars
```{r}
hyp1Dat <- hyp1Dat %>%
  mutate_at(vars(negPanas, posPanas), ~na_if(., 8))

hyp2Dat <- hyp2Dat %>%
  #filter(dailyMeanNegAff <= 4)
  mutate_at(vars(contains("daily")), ~na_if(., 7)) %>%
  mutate_at(vars(contains("daily")), ~na_if(., 8)) %>%
  mutate_at(vars(contains("daily")), ~na_if(., 9))

#need to assign na for rest of vars
```

# Summary of hypothesis 1 data

## Summary of factor variables
```{r, message=F, warning=F, collapse=T}
# convert to factors
hyp1Dat <- hyp1Dat %>%
  mutate(gender = factor(gender, levels = c(1, 2), 
                         labels = c("male", "female"))) %>%
  mutate(medUseNone1Yes = factor(medUseNone1Yes, levels = c(1, 2, 8),
                                 labels = c("yes", "no", "missing"))) %>%
  mutate(ethnicity = factor(ethnicity, levels = c(1, 2, 3, 4, 6, 7, 8),
                            labels = c("caucasian", "africanAmerican",
                                       "nativeAmerican", "asian", "other",
                                       "dontKnow", "refused"))) %>%
  mutate(education = factor(education, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9,
                                                  10, 11, 12, 97),
                            labels = c("grades16", "grades78", "grades\n912NoGed",
                                       "ged", "hsGrad", "college\n12YrNoDegree",
                                       "college\n3OrMore\nYrNoDegree",
                                       "college\n2YrGrad", "college\n4YrGrad",
                                       "some\nGradSchl", "master\nDegree",
                                       "doctoral\nDegree", "dontKnow"))) %>%
  mutate(smoker = factor(smoker, levels = c(1, 2, 9), 
                         labels = c("yes", "no", "inapp")))
```

### Gender
```{r, message=F, warning=F, collapse=T}
# gender
hyp1Dat %>%
  group_by(gender) %>%
  ggplot(., aes(gender)) +
    geom_bar(aes(fill = gender)) +
    geom_text(stat='count', aes(label = scales::percent(round(prop.table(stat(count)), 3))), vjust=-.50, 
              color = "gray29", size = 3)
```

### Medication Use
```{r, message=F, warning=F, collapse=T}
# medication use
hyp1Dat %>%
  group_by(medUseNone1Yes) %>%
  ggplot(., aes(medUseNone1Yes)) +
    geom_bar(aes(fill = medUseNone1Yes)) +
    geom_text(stat='count', aes(label = scales::percent(round(prop.table(stat(count)), 4))), vjust=-.50, 
              color = "gray29", size = 3)
```

### Ethnicity
```{r, message=F, warning=F, collapse=T}
# ethnicity
hyp1Dat %>%
  group_by(ethnicity) %>%
  ggplot(., aes(ethnicity)) +
    geom_bar(aes(fill = ethnicity)) +
    geom_text(stat='count', aes(label = scales::percent(round(prop.table(stat(count)), 3))), vjust=-.50, 
              color = "gray29", size = 3) +
    theme(axis.text.x = element_text(angle = 10))
```

### Education
```{r, message=F, warning=F, collapse=T}
# education
hyp1Dat %>%
  mutate(education = factor(education, levels = c("hsGrad", "college\n4YrGrad",
                                       "college\n12YrNoDegree", "master\nDegree",
                                       "NA", "college\n2YrGrad", "doctoral\nDegree",
                                       "college\n3OrMore\nYrNoDegree",
                                       "some\nGradSchl", "grades\n912NoGed",
                                       "grades78", "ged", "grades16", 
                                       "dontKnow"))) %>%
  group_by(education) %>%
  ggplot(., aes(education)) +
    geom_bar(aes(fill = education)) +
    geom_text(stat='count', aes(label = scales::percent(round(prop.table(stat(count)), 4))), vjust=-.50, 
              color = "gray29", size = 2.5) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylim(0, 400)
```

### Smoker Status
```{r, message=F, warning=F, collapse=T}
# smoker
ggplot(hyp1Dat, aes(smoker)) +
  geom_bar(aes(fill = smoker)) +
  geom_text(stat='count', aes(label = scales::percent(prop.table(stat(count)))), vjust=-.50, 
              color = "gray29") 
```

# Summary of numeric variables
```{r, , message=F, warning=F, attr.output='style="max-height: 500px;"'}
hyp1Dat %>%
  summarise_if(is.numeric,
               funs(sum(is.na(.)))) %>%
  pivot_longer(1:39, names_to = "summaryStats", values_to = "numberOfNAs") %>% print(n = Inf)

hyp1Dat %>%
  summarise_if(is.numeric,
               lst(mean, median, sd, min, max),
               na.rm = T) %>%
  pivot_longer(1:195, names_to = "summaryStats", values_to = "values") %>% 
  print(n = Inf) %>%
  separate(summaryStats, c("variable", "stat"), "_") %>%
  arrange(variable)
```

### Correlation matrix of cortisol variables
```{r, message=F, warning=F, collapse=T, fig.height=9}
hyp1Dat %>%
  select(car, aucTot, lmmCortSlope, hyp1LmmUncCiiv, hyp1LmmCondCiiv) %>%
  GGally::ggpairs(.)
```

### Histogram of LMM cIIV, conditional on time
```{r, message=F, warning=F, collapse=T}
ggplot(hyp1Dat, aes(hyp1LmmCondCiiv)) +
  geom_histogram(binwidth = .1)
```

### Histogram of positive and negative PANAS
```{r, message=F, warning=F, collapse=T}
ggplot(hyp1Dat, aes(negPanas)) +
  geom_histogram(binwidth = .3)

ggplot(hyp1Dat, aes(posPanas)) +
  geom_histogram(binwidth = .3)
```

# Summary of hypothesis 2 data
```{r, message=F, warning=F, collapse=T}
hyp2Dat %>%
  group_by(day) %>%
  summarise(meanPosAffectPerDay = mean(dailyMeanPosAff, na.rm = T),
            meanNegAffectPerDay = mean(dailyMeanNegAff, na.rm = T))
```

## Histograms of day 2 positive and negative affect
```{r, message=F, warning=F, collapse=T}
hyp2Dat %>%
  filter(day == 2) %>%
  ggplot(aes(dailyMeanPosAff)) +
  geom_histogram(binwidth = .3)

hyp2Dat %>%
  filter(day == 2) %>%
  ggplot(aes(dailyMeanNegAff)) +
  geom_histogram(binwidth = .1)
```

## Histograms of day 3 positive and negative affect
```{r, message=F, warning=F, collapse=T}
hyp2Dat %>%
  filter(day == 3) %>%
  ggplot(aes(dailyMeanPosAff)) +
  geom_histogram(binwidth = .3)

hyp2Dat %>%
  filter(day == 3) %>%
  ggplot(aes(dailyMeanNegAff)) +
  geom_histogram(binwidth = .1)
```

## Histograms of day 4 positive and negative affect
```{r, message=F, warning=F, collapse=T}
hyp2Dat %>%
  filter(day == 4) %>%
  ggplot(aes(dailyMeanPosAff)) +
  geom_histogram(binwidth = .1)

hyp2Dat %>%
  filter(day == 4) %>%
  ggplot(aes(dailyMeanNegAff)) +
  geom_histogram(binwidth = .1)
```

## Histograms of day 4 positive and negative affect
```{r, message=F, warning=F, collapse=T}
hyp2Dat %>%
  filter(day == 4) %>%
  ggplot(aes(dailyMeanPosAff)) +
  geom_histogram(binwidth = .1)

hyp2Dat %>%
  filter(day == 4) %>%
  ggplot(aes(dailyMeanNegAff)) +
  geom_histogram(binwidth = .1)
```

## Histograms of day 5 positive and negative affect
```{r, message=F, warning=F, collapse=T}
hyp2Dat %>%
  filter(day == 5) %>%
  ggplot(aes(dailyMeanPosAff)) +
  geom_histogram(binwidth = .1)

hyp2Dat %>%
  filter(day == 5) %>%
  ggplot(aes(dailyMeanNegAff)) +
  geom_histogram(binwidth = .1)
```


























