---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

# Load libraries
```{r, message=F, warning=F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(nlme)
options(scipen=999)
```

# Load daily data
```{r, message=F, warning=F, results="hold"}
dailyDat <- read.table("/Users/Santi/Desktop/rProjects/dissertation/analysis/data.tsv",
                  header = T,
                  sep = "\t",
                  stringsAsFactors = F,
                  na.strings = c(999994, 999995, 999996, 999998, 98.00))

# 999994. UNRELIABLE
# 999995. NOT DONE
# 999996. EMPTY
# 999998. REFUSED/MISSING
```

# Select cortisol variables of interest from daily data
```{r, message=F, warning=F, results="hold"}
cortTimesDat <- dailyDat %>%
  select(id = M2ID, gender = B1PGENDER, age = B1PAGE_M2, medUseNone1Yes = B2DMED10, startWork = B2DA3AAP, endWork = B2DA3BAP, providedCort1yes2no = B2DCORT, day = B2DDAY, wakeCort = B2DCORWT, wakeThirtyCort = B2DCORAT, noonCort = B2DCORLT, pmCort = B2DCORBT) %>%
  filter(providedCort1yes2no == 1) %>%
  filter(day == 2 | day == 3 | day == 4 | day == 5) %>%
  pivot_longer(9:12, names_to = "cortTimeText", values_to = "cortTime") 
  
cortValuesDat <- dailyDat %>%
  select(id = M2ID, providedCort1yes2no = B2DCORT, day = B2DDAY, wakeCort = B2DCORW, wakeThirtyCort = B2DCORA, noonCort = B2DCORL, pmCort = B2DCORB) %>%
  filter(providedCort1yes2no == 1) %>%
  filter(day == 2 | day == 3 | day == 4 | day == 5) %>%
  pivot_longer(4:7, names_to = "cortTimeText", values_to = "cortNmol") %>%
  select(id, day, cortTimeText, cortNmol)

cortDat <- left_join(cortTimesDat, cortValuesDat, by = c("id", "day", "cortTimeText"))

str(cortDat)
```

# Remove NAs from cortTime and cortNmol
```{r, message=F, warning=F, results="hold"}
naCortNmol <- round((sum(is.na(cortDat$cortNmol)) / nrow(cortDat)) * 100, 2)
naCortTime <- round((sum(is.na(cortDat$cortTime)) / nrow(cortDat)) * 100, 2)

masterCortDat <- cortDat %>%
  filter(!is.na(cortNmol), !is.na(cortTime))

str(masterCortDat)
```

# Remove samples collected outside of 24hr window
```{r, message=F, warning=F, results="hold"}
samplesOutside24hrWindow <- masterCortDat %>%
  filter(cortTime < 0.00 | cortTime > 24.00) %>%
  nrow()
#607 samples collected prior to beginning of of interview day or after end of interview day

masterCortDat <- masterCortDat %>%
  filter(cortTime >= 0.00, cortTime <= 24.00)

str(masterCortDat)
```

# Remove samples with cortNmol > 60
```{r, message=F, warning=F, results="hold"}
masterCortDat %>%
  filter(cortNmol > 60)
#381 greater than 60

greaterThanSixty <- round((nrow(subset(masterCortDat, cortNmol > 60)) / nrow(masterCortDat)) * 100, 2)

masterCortDat <- masterCortDat %>%
  filter(cortNmol <= 60)

str(masterCortDat)
```

# Remove waking cortisol samples collected before 0400h or after 1200h
```{r, message=F, warning=F, results="hold"}
earlyLateWakeCort <- masterCortDat %>%
  filter(cortTimeText == "wakeCort") %>%
  filter(cortTime < 4.00 | cortTime > 12) %>%
  nrow()
#173 wake cort collected before 4am or after 12pm

# df that contains wakeCortTime between 4 and 12
wakeCortKeepMe <- masterCortDat %>%
  filter(cortTimeText == "wakeCort") %>%
  filter(between(cortTime, 4.00, 12.00)) 

# remove all wakeCort rows from master dat
masterCortDat <- masterCortDat %>%
  filter(cortTimeText != "wakeCort")

# bind wakeCortKeepMe rows w/ masterCortDat rows
masterCortDat <- bind_rows(masterCortDat, wakeCortKeepMe)

# arrange rows
masterCortDat <- masterCortDat %>%
  arrange(id, day, cortTime)

str(masterCortDat)
```

# Number of 1200h Values that are 10 nmol/L greater than the wake + 30 values?
```{r, message=F, warning=F, results="hold"}
erronLunchValues <- masterCortDat %>% 
  mutate(dummy1 = case_when(
    cortTimeText == "wakeThirtyCort" ~ 1,
    cortTimeText == "noonCort" ~ 1,
    TRUE ~ 0
  )) %>%
  filter(dummy1 == 1) %>%
  arrange(id, day, cortTime) %>%
  group_by(id, day) %>%
  mutate(dummyDif = last(cortNmol) - first(cortNmol)) %>%
  mutate(assignDummy = case_when(
    dummyDif > 10 ~ 1,
    TRUE ~ 0
  )) %>%
  filter(assignDummy == 1) %>%
  nrow()
# 166 samples

nmolWakeNoon10 <- masterCortDat %>%
  mutate(dummy1 = case_when(
    cortTimeText == "wakeThirtyCort" ~ 1,
    cortTimeText == "noonCort" ~ 1,
    TRUE ~ 0
  )) %>%
  filter(dummy1 == 1) %>%
  arrange(id, day, cortTime) %>%
  group_by(id, day) %>%
  mutate(dummyDif = last(cortNmol) - first(cortNmol)) %>%
  mutate(assignDummy = case_when(
    dummyDif > 10 ~ 1,
    TRUE ~ 0
  )) %>%
  filter(assignDummy == 1) %>%
  select(-dummy1, -dummyDif, -assignDummy)

masterCortDat <- anti_join(masterCortDat, nmolWakeNoon10)

str(masterCortDat)
```

# How many participants are using medications that impact HPA axis function,
#didn't know or did not answer this question?
```{r, message=F, warning=F, results="hold"}
masterCortDat %>%
  filter(medUseNone1Yes == 2 | medUseNone1Yes == 8 | medUseNone1Yes == 9) %>%
  select(id) %>%
  distinct() %>%
  nrow()
# 895 of 1731

masterCortDat %>%
  select(id) %>%
  n_distinct()
```

# Compute natural logarithm of cortisol
```{r, message=F, warning=F, results="hold"}
masterCortDat <- masterCortDat %>%
  mutate(logCort = log(cortNmol))

str(masterCortDat)
```

# Derive CAR
```{r, message=F, warning=F, results="hold"}
masterCortDat <- masterCortDat %>%
  group_by(id) %>%
  mutate(car = nth(logCort, 2) - nth(logCort, 1)) %>%
  ungroup()

str(masterCortDat)
```

# Derive AUC
```{r, message=F, warning=F, results="hold"}
aucDat <- masterCortDat %>%
  select(id, day, cortTimeText, cortTime, logCort) %>%
  filter(cortTimeText != "wakeThirtyCort") %>%
  group_by(id, cortTimeText) %>%
  mutate(meanCortTime = mean(cortTime)) %>%
  mutate(meanLogCort = mean(logCort)) %>%
  ungroup() %>%
  filter(day == 2) %>%
  group_by(id) %>%
  mutate(duration = meanCortTime - lag(meanCortTime)) %>%
  mutate(auc = ((meanLogCort + lag(meanLogCort)) / 2) * duration ) %>%
  mutate(aucTot = sum(auc, na.rm = T)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(any(n() == 3)) %>%
  select(id, aucTot) %>%
  distinct() 

# check who has negative AUC and why??? aha - remove any(id) groups where n(timepoints) < 3
aucLessThanTimePoints <- masterCortDat %>%
  select(id, day, cortTimeText, cortTime, logCort) %>%
  filter(cortTimeText != "wakeThirtyCort") %>%
  group_by(id, cortTimeText) %>%
  mutate(meanCortTime = mean(cortTime)) %>%
  mutate(meanLogCort = mean(logCort)) %>%
  ungroup() %>%
  filter(day == 2) %>%
  group_by(id) %>%
  mutate(duration = meanCortTime - lag(meanCortTime)) %>%
  mutate(auc = ((meanLogCort + lag(meanLogCort)) / 2) * duration ) %>%
  mutate(aucTot = sum(auc, na.rm = T)) %>%
  ungroup() %>%
  group_by(id) %>%
  filter(any(n() < 3)) %>%
  select(id) %>%
  distinct() %>%
  ungroup()

masterCortDat <- left_join(masterCortDat, aucDat, by = "id")

str(masterCortDat)
```

# Derive slope
```{r, message=F, warning=F, results="hold"}
cortSlopeDat <- masterCortDat %>%
  select(id, day, cortTimeText, cortTime, logCort) %>%
  filter(cortTimeText != "wakeThirtyCort")


lmeSlopes <- gls(logCort ~ 1,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude")

lmeSlopes2 <- lme(logCort ~ cortTime,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 random = ~1|id)

lmeSlopes3 <- lme(logCort ~ cortTime,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 random = ~cortTime|id)

anova(lmeSlopes, lmeSlopes2, lmeSlopes3)

slopeDat <- as.data.frame(coef(lmeSlopes3))

slopeDat <- slopeDat %>%
  tibble::rownames_to_column("id") %>%
  rename(lmmCortSlope = cortTime) %>%
  select(-"(Intercept)") %>%
  mutate(lmmCortSlope = round(lmmCortSlope, 2)) %>%
  mutate(id = as.numeric(id))

indRegSlope <- cortSlopeDat %>%
  group_by(id) %>% 
  do(broom::tidy(lm(logCort ~ cortTime, data=.,
             na.action = na.exclude))) %>% #change predictor 
  filter(., term == "cortTime") %>%
  dplyr::select(., id, leastSquaresCortSlope = estimate) %>%
  mutate(leastSquaresCortSlope = round(leastSquaresCortSlope, 2))

tTestJoin <- left_join(indRegSlope, slopeDat, by = "id")

cor.test(tTestJoin$lmmCortSlope, tTestJoin$leastSquaresCortSlope,
         paired = T) #.7869

masterCortDat <- left_join(masterCortDat, slopeDat, by = "id")
masterCortDat <- left_join(masterCortDat, indRegSlope, by = "id")

str(masterCortDat)
```

# NAs per slope derivation method
```{r, message=F, warning=F, results="hold"}
masterCortDat %>%
  summarise(sumLmmNA = sum(is.na(lmmCortSlope)),
            sumLsNA = sum(is.na(leastSquaresCortSlope)))
```

# LMM and least-squares association with observed values: all days 
```{r, message=F, warning=F, results="hold"}
#extract observed
observedValues <- cortSlopeDat %>%
  select(id, logCort) %>%
  na.omit()

#extract fitted lme
fittedLmeValues <- fitted(lmeSlopes3) %>%
  tibble::enframe(., "id", "fittedLme")

#extract fitted least-squares
fittedLeastSquaresValues <- cortSlopeDat %>%
  group_by(id) %>% 
  do(broom::tidy(lm(logCort ~ cortTime, data=.,
             na.action = na.exclude)$fitted)) %>% #change predictor 
  select(id, fittedLeastSquares = x)

#compare
cor.test(observedValues$logCort, fittedLmeValues$fittedLme, 
         method = "spearman",
         paired = T) 

cor.test(observedValues$logCort, fittedLeastSquaresValues$fittedLeastSquares, 
         method = "spearman",
         paired = T)
```
# LMM and least-squares association with observed values: each day, cond only
```{r, message=F, warning=F, results="hold"}
#extract observed
observedValues <- cortSlopeDat %>%
  dplyr::select(id, day, logCort) %>%
  na.omit()

##lme approach separate days: day 2
lmeFitDay2Dat <- cortSlopeDat %>%
  filter(day == 2)

lmeResidsDay2 <- lme(logCort ~ cortTime,
                 data = lmeFitDay2Dat,
                 method = "ML",
                 na.action = "na.exclude",
                 control = list(opt = "optim"),
                 random = ~cortTime|id)

#extract fitted lme
fittedLmeValuesDay2 <- fitted(lmeResidsDay2) %>%
  tibble::enframe(., "id", "fittedLmeDay2") %>%
  mutate(id = as.numeric(id))

attr(fittedLmeValuesDay2$fittedLmeDay2, "label") <- NULL

#cor result
observedValuesDay2 <- observedValues %>%
  filter(day == 2)

cor.test(fittedLmeValuesDay2$fittedLmeDay2, observedValuesDay2$logCort, 
         method = "spearman",
         paired = T)

##lme approach separate days: day 3
lmeFitDay3Dat <- cortSlopeDat %>%
  filter(day == 3)

lmeResidsDay3 <- lme(logCort ~ cortTime,
                 data = lmeFitDay3Dat,
                 method = "ML",
                 na.action = "na.exclude",
                 control = list(opt = "optim"),
                 random = ~cortTime|id)

#extract fitted lme
fittedLmeValuesDay3 <- fitted(lmeResidsDay3) %>%
  tibble::enframe(., "id", "fittedLmeDay3") %>%
  mutate(id = as.numeric(id))

attr(fittedLmeValuesDay3$fittedLmeDay3, "label") <- NULL

#cor result
observedValuesDay3 <- observedValues %>%
  filter(day == 3)

cor.test(fittedLmeValuesDay3$fittedLmeDay3, observedValuesDay3$logCort, 
         method = "spearman",
         paired = T)

##lme approach separate days: day 4
lmeFitDay4Dat <- cortSlopeDat %>%
  filter(day == 4)

lmeResidsDay4 <- lme(logCort ~ cortTime,
                 data = lmeFitDay4Dat,
                 method = "ML",
                 na.action = "na.exclude",
                 control = list(opt = "optim"),
                 random = ~cortTime|id)

#extract fitted lme
fittedLmeValuesDay4 <- fitted(lmeResidsDay4) %>%
  tibble::enframe(., "id", "fittedLmeDay4") %>%
  mutate(id = as.numeric(id))

attr(fittedLmeValuesDay4$fittedLmeDay4, "label") <- NULL

#cor result
observedValuesDay4 <- observedValues %>%
  filter(day == 4)

cor.test(fittedLmeValuesDay4$fittedLmeDay4, observedValuesDay4$logCort, 
         method = "spearman",
         paired = T)

##lme approach separate days: day 5
lmeFitDay5Dat <- cortSlopeDat %>%
  filter(day == 5)

lmeResidsDay5 <- lme(logCort ~ cortTime,
                 data = lmeFitDay5Dat,
                 method = "ML",
                 na.action = "na.exclude",
                 control = list(opt = "optim"),
                 random = ~cortTime|id)

#extract fitted lme
fittedLmeValuesDay5 <- fitted(lmeResidsDay5) %>%
  tibble::enframe(., "id", "fittedLmeDay5") %>%
  mutate(id = as.numeric(id))

attr(fittedLmeValuesDay5$fittedLmeDay5, "label") <- NULL

#cor result
observedValuesDay5 <- observedValues %>%
  filter(day == 5)

cor.test(fittedLmeValuesDay5$fittedLmeDay5, observedValuesDay5$logCort, 
         method = "spearman",
         paired = T)

##lme approach with days nested, cond on time
lmeResids <- lme(logCort ~ cortTime,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 control = list(opt = "optim"),
                 random = ~cortTime|id/day)

#extract fitted lme
fittedLmeValues <- fitted(lmeResids) %>%
  tibble::enframe(., "id", "fittedLme")

fittedLmeValues <- fittedLmeValues %>%
  separate(id, c("id", "day"), "/") %>%
  mutate(id = as.numeric(id), day = as.numeric(day))

attr(fittedLmeValues$fittedLme, "label") <- NULL

fitted3LevelDay2 <- fittedLmeValues %>%
  filter(day == 2)

cor.test(fitted3LevelDay2$fittedLme, observedValuesDay2$logCort, 
         method = "spearman",
         paired = T)

fitted3LevelDay3 <- fittedLmeValues %>%
  filter(day == 3)

cor.test(fitted3LevelDay3$fittedLme, observedValuesDay3$logCort, 
         method = "spearman",
         paired = T)

fitted4LevelDay4 <- fittedLmeValues %>%
  filter(day == 4)

cor.test(fitted4LevelDay4$fittedLme, observedValuesDay4$logCort, 
         method = "spearman",
         paired = T)

fitted5LevelDay5 <- fittedLmeValues %>%
  filter(day == 5)

cor.test(fitted5LevelDay5$fittedLme, observedValuesDay5$logCort, 
         method = "spearman",
         paired = T)

#cond ciiv least squares
lsFittedHyp2Cond <- cortSlopeDat %>% 
  group_by(id, day) %>%
  add_count(name = "numberOfValuesPerIdDay") %>%
  do(broom::tidy(lm(logCort ~ cortTime, data=., 
             na.action = na.exclude)$fitted)) %>%
  select(id, day, fittedLeastSquares = x) %>%
  ungroup()

lsFittedHyp2CondDay2 <- lsFittedHyp2Cond %>%
  filter(day == 2)

lsFittedHyp2CondDay3 <- lsFittedHyp2Cond %>%
  filter(day == 3)

lsFittedHyp2CondDay4 <- lsFittedHyp2Cond %>%
  filter(day == 4)

lsFittedHyp2CondDay5 <- lsFittedHyp2Cond %>%
  filter(day == 5)

cor.test(lsFittedHyp2CondDay2$fittedLeastSquares, observedValuesDay2$logCort, 
         method = "spearman",
         paired = T)
cor.test(lsFittedHyp2CondDay3$fittedLeastSquares, observedValuesDay3$logCort, 
         method = "spearman",
         paired = T)
cor.test(lsFittedHyp2CondDay4$fittedLeastSquares, observedValuesDay4$logCort, 
         method = "spearman",
         paired = T)
cor.test(lsFittedHyp2CondDay5$fittedLeastSquares, observedValuesDay5$logCort, 
         method = "spearman",
         paired = T)
```

# Hypotheses 1a and 1b cIIV derivation: unconditional and conditional on time
```{r, message=F, warning=F, results="hold"}
#hyp 1 lmm uncond ciiv
lmeSlopes <- gls(logCort ~ 1,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude")

lmeSlopes2 <- lme(logCort ~ 1,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 random = ~1|id)

anova(lmeSlopes, lmeSlopes2)

lmeResidsHyp1Unc <- resid(lmeSlopes2, type="response", level=1)
lmeResidsHyp1Unc <- tibble::enframe(lmeResidsHyp1Unc, "id", "hyp1LmmUncCiiv")

lmeResidsHyp1Unc <- lmeResidsHyp1Unc %>%
  group_by(id) %>%
  mutate(hyp1LmmUncCiiv = sd(hyp1LmmUncCiiv, na.rm = T)) %>%
  ungroup() %>%
  select(id, hyp1LmmUncCiiv) %>%
  distinct() %>%
  mutate(id = as.numeric(id))

masterCortDat <- left_join(masterCortDat, lmeResidsHyp1Unc, by = "id")

#hyp 1 lmm cond ciiv
lmeSlopes2 <- lme(logCort ~ 1,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 random = ~1|id)

lmeSlopes3 <- lme(logCort ~ cortTime,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 random = ~1|id)

lmeSlopes4 <- lme(logCort ~ cortTime,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 random = ~cortTime|id)

anova(lmeSlopes2, lmeSlopes3, lmeSlopes4)

lmeResidsHyp1Cond <- resid(lmeSlopes4, type="response", level=1)
lmeResidsHyp1Cond <- tibble::enframe(lmeResidsHyp1Cond, "id", "hyp1LmmCondCiiv")

lmeResidsHyp1Cond <- lmeResidsHyp1Cond %>%
  group_by(id) %>%
  mutate(hyp1LmmCondCiiv = sd(hyp1LmmCondCiiv, na.rm = T)) %>%
  ungroup() %>%
  select(id, hyp1LmmCondCiiv) %>%
  distinct() %>%
  mutate(id = as.numeric(id))

masterCortDat <- left_join(masterCortDat, lmeResidsHyp1Cond, by = "id")

#hyp 1 unc ciiv least squares / not actually least-squares, just sd ## CHANGE NAME
sdResidsHyp1Unc <- cortSlopeDat %>%
  group_by(id) %>%
  mutate(hyp1SdUncCiiv = sd(logCort, na.rm = T)) %>%
  ungroup() %>%
  select(id, hyp1SdUncCiiv) %>%
  distinct()

masterCortDat <- left_join(masterCortDat, sdResidsHyp1Unc, by = "id")

# hyp1 cond ciiv least squares
lsResidsHyp1Cond <- cortSlopeDat %>%
  group_by(id) %>%
  do(broom::tidy(lm(logCort ~ cortTime, data=., 
             na.action = na.exclude)$residuals)) %>% #change predictor
  mutate(hyp1LsCondCiiv = sd(x, na.rm = T)) %>%
  select(., id, hyp1LsCondCiiv) %>%
  distinct()

masterCortDat <- left_join(masterCortDat, lsResidsHyp1Cond, by = "id")

#cor 
masterCortDat %>%
  select(id, hyp1SdUncCiiv, hyp1LmmUncCiiv) %>%
  distinct() %>%
  na.omit() %>%
  do(broom::tidy(cor.test(.$hyp1SdUncCiiv, .$hyp1LmmUncCiiv, 
                          method = "spearman", paired = T)))

masterCortDat %>%
  select(id, hyp1LsCondCiiv, hyp1LmmCondCiiv) %>%
  distinct() %>%
  na.omit() %>%
  do(broom::tidy(cor.test(.$hyp1LsCondCiiv, .$hyp1LmmCondCiiv, 
                          method = "spearman", paired = T)))
```

# Hypotheses 2a and 2b cIIV derivation: unconditional and conditional on time
```{r, message=F, warning=F, results="hold"}
#lme approach, unconditional
lmeResids <- lme(logCort ~ 1,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 random = ~1|id/day)

lmeResids <- resid(lmeResids, type="response", level=2)
lmeResids <- tibble::enframe(lmeResids, "id", "cIiv") # name this resid something

lmeResids <- lmeResids %>%
  separate(id, c("id", "day"), "/") 

lmeResidsHyp2Unc <- lmeResids %>%
  group_by(id, day) %>%
  mutate(hyp2LmmUncCiiv = sd(cIiv, na.rm = T)) %>%
  ungroup() %>%
  select(id, day, hyp2LmmUncCiiv) %>%
  distinct() %>%
  mutate(id = as.numeric(id), day = as.numeric(day))

masterCortDat <- left_join(masterCortDat, lmeResidsHyp2Unc, by = c("id", "day"))

##lme approach with days nested, cond on time
lmeResids <- lme(logCort ~ cortTime,
                 data = cortSlopeDat,
                 method = "ML",
                 na.action = "na.exclude",
                 random = ~1|id/day)

lmeResids <- resid(lmeResids, type="response", level=2)
lmeResids <- tibble::enframe(lmeResids, "id", "cIiv") # name this resid something

lmeResids <- lmeResids %>%
  separate(id, c("id", "day"), "/") 

lmeResidsHyp2Cond <- lmeResids %>%
  group_by(id, day) %>%
  mutate(hyp2LmmCondCiiv = sd(cIiv, na.rm = T)) %>%
  ungroup() %>%
  select(id, day, hyp2LmmCondCiiv) %>%
  distinct() %>%
  mutate(id = as.numeric(id), day = as.numeric(day))

masterCortDat <- left_join(masterCortDat, lmeResidsHyp2Cond, by = c("id", "day"))

#unconditional ciiv least squares
sdResidsHyp2Unc <- cortSlopeDat %>%
  group_by(id, day) %>%
  mutate(hyp2SdUncCiiv = sd(logCort)) %>%
  ungroup() %>%
  select(id, day, hyp2SdUncCiiv) %>%
  distinct()

masterCortDat <- left_join(masterCortDat, sdResidsHyp2Unc, by = c("id", "day"))

#cond ciiv least squares
lsResidsHyp2Cond <- cortSlopeDat %>%
  group_by(id, day) %>%
  add_count(name = "numberOfValuesPerIdDay") %>%
  do(broom::tidy(lm(logCort ~ cortTime, data=., 
             na.action = na.exclude)$residuals)) %>% 
  mutate(hyp2LsCondCiiv = sd(x, na.rm = T)) %>%
  ungroup() %>%
  select(., id, day, hyp2LsCondCiiv) %>%
  distinct()

masterCortDat <- left_join(masterCortDat, lsResidsHyp2Cond, by = c("id", "day"))

# NAs per slope derivation method
masterCortDat %>%
  summarise(sumLmmNA = sum(is.na(hyp2LmmCondCiiv)),
            sumLsNA = sum(is.na(hyp2LsCondCiiv)))

str(masterCortDat)
```

# Select cortisol variables for each hypothesis and remove duplicates
```{r, message=F, warning=F, results="hold"}
hyp1CortDat <- masterCortDat %>%
  select(id, age, gender, medUseNone1Yes, car, aucTot, lmmCortSlope,
         leastSquaresCortSlope, hyp1LmmUncCiiv, hyp1LmmCondCiiv,
         hyp1SdUncCiiv, hyp1LsCondCiiv) %>%
  distinct()

str(hyp1CortDat)

hyp2CortDat <- masterCortDat %>%
  select(id, age, gender, medUseNone1Yes, day, hyp2LmmUncCiiv,
         hyp2LmmCondCiiv, hyp2SdUncCiiv, hyp2LsCondCiiv) %>%
  distinct()

str(hyp2CortDat)
```

# Select emotion data and join with hypotheses 1 & 2 cortisol data
```{r, message=F, warning=F, results="hold"}
emoDat <- dailyDat %>%
  select(id = M2ID, day = B2DDAY, dailyMeanPosAff = B2DPOSAV, dailyMeanNegAffect = B2DNEGAV, dailyInGoodSpirits = B2DC7, dailyCheerful = B2DC8, dailyExtremelyHappy = B2DC9, dailyCalmPeaceful = B2DC10, dailySatisfied = B2DC11, dailyFullOfLife = B2DC12, dailyCloseToOthers = B2DC21, dailyLikeYouBelong = B2DC22, dailyEnthusiastic = B2DC23, dailyAttentive = B2DC24, dailyProud = B2DC25, dailyActive = B2DC26, dailyConfident = B2DC27, dailyRestlessFidgety = B2DC1, dailyNervous = B2DC2, dailyWorthless = B2DC3, dailySoSadNoCheerUp = B2DC4, dailyEverythingAnEffort = B2DC5, dailyHopeless = B2DC6, dailyLonely = B2DC13, dailyAfraid = B2DC14, dailyJittery = B2DC15, dailyIrritable = B2DC16, dailyAshamed = B2DC17, dailyUpset = B2DC18, dailyAngry = B2DC19, dailyFrustrated = B2DC20) %>%
  filter(day == 2 | day == 3 | day == 4 | day == 5)
```

# Select variables of interest from midus 2 data
```{r, message=F, warning=F, results="hold"}
midusDat <- read.table("/Users/Santi/Desktop/rProjects/dissertation/analysis/midusTwoData.tsv",
                  header = T,
                  sep = "\t",
                  stringsAsFactors = F)

midusDat <- midusDat %>%
  select(id = M2ID, ethnicity = B1PF7A, education = B1PB1, smoker = B1PA39, posPanas = B1SPOSPA, negPanas = B1SNEGPA, traitCheerful = B1SA26A, traitGoodSpirits = B1SA26B, traitExtremelyHappy = B1SA26C, traitCalmPeaceful = B1SA26D, traitSatisfied = B1SA26E, traitFullOfLife = B1SA26F, traitCloseOthers = B1SA26G, traitBelong = B1SA26H, traitEnthusiastic = B1SA26I, traitAttentive = B1SA26J, traitProud = B1SA26K, traitActive = B1SA26L, traitConfident = B1SA26M, traitSoSadNoCheer = B1SA24A, traitNervous = B1SA24B, traitRestless = B1SA24C, traitHopeless = B1SA24D, traitEverythingAnEffort = B1SA24E, traitWorthless = B1SA24F, traitLonely = B1SA24G, traitAfraid = B1SA24H, traitJittery = B1SA24I, traitIrritable = B1SA24J, traitAshamed = B1SA24K, traitUpset = B1SA24L, traitAngry = B1SA24M, traitFrustrated = B1SA24N)
```

# Hypothesis 1 dataframe
```{r, message=F, warning=F, results="hold"}
hyp1Dat <- left_join(hyp1CortDat, midusDat, by = "id") %>%
  select(id, age, gender, medUseNone1Yes, ethnicity, 
         education, smoker, everything())

str(hyp1Dat)
hyp1Dat
write.csv(hyp1Dat, "hyp1Dat.csv", row.names = F)
```

# Hypothesis 2 dataframe
```{r, message=F, warning=F, results="hold"}
hyp2Dat <- left_join(hyp2CortDat, emoDat, by = c("id", "day"))

str(hyp2Dat)
hyp2Dat
write.csv(hyp2Dat, "hyp2Dat.csv", row.names = F)
```








































