---
output:
  pdf_document:
    fig_caption: yes
    number_sections: no
  word_document: default
bibliography: refs.bib
csl: apa-6th-edition.csl
geometry: margin = 1.0in
font-family: Times New Roman
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{sectsty}
- \allsectionsfont{\centering\normalsize}
- \subsectionfont{\centering\normalsize}
- \subsubsectionfont{\raggedright\normalsize}
- \usepackage{titlesec}
- \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \usepackage[document]{ragged2e}
- \raggedright
- \setlength{\parindent}{4em}
- \setlength{\parskip}{0em}
- \setlength{\RaggedRightParindent}{0.5in}
---

```{r setup, include = FALSE}
library("papaja")
library("rmarkdown")
#table of contents: https://stackoverflow.com/questions/23957278/how-to-add-table-of-contents-in-rmarkdown

# indent: https://latex.org/forum/viewtopic.php?t=28495

#rmarkdown::render("analysis/manuscript.Rmd", rmarkdown::pdf_document())
```

\setcounter{page}{1}

\begin{center}
MY TITLE\\
\vspace{4cm}
A Dissertation\\
Presented to the Faculty of\\
The PGSP-Stanford Psy.D. Consortium\\
Palo Alto University\\
Palo Alto, California\\
In Partial Fulfillment of the\\
Requirements for the Degree of\\
Doctor of Psychology\\
\vspace{8.5cm}
by\\
Santiago Allende\\
November, 2019
\end{center}

\newpage
\begin{center}
MY TITLE\\
Santiago Allende\\
The PGSP-Stanford Psy.D. Consortium, Palo Alto University, 2019
\end{center}

\newpage
\vspace*{\fill}
\begingroup
\centering

\textcopyright\space 2019\\
by\\
Santiago Allende\\
 
\endgroup
\vspace*{\fill}

\newpage
\begin{center}
My Title\\
\vspace{3cm}
This dissertation by Santiago Allende, directed and approved by the candidate’s committee, has\\
been accepted and approved by the Faculty of The PGSP-Stanford Psy.D. Consortium, Palo Alto\\
University in partial fulfillment of the requirements for the degree of\\
DOCTOR OF PSYCHOLOGY\\
\vspace{3cm}
\begin{minipage}{4cm}
  \centering
  \hrulefill\\
  Signed
\end{minipage}

\end{center}

\newpage

\newpage
\tableofcontents 
\listoffigures
\listoftables
\newpage

```{r 'chapter1.Rmd', child = "/Users/Santi/Desktop/rProjects/dissertation/analysis/chapter1.Rmd"}
```

\newpage

```{r 'hypotheses.Rmd', child = "/Users/Santi/Desktop/rProjects/dissertation/analysis/hypotheses.Rmd"}
```

\newpage

```{r 'chapter2.Rmd', child = "/Users/Santi/Desktop/rProjects/dissertation/analysis/chapter2.Rmd"}
```


















\newpage

# References
```{r create_r-references}
r_refs(file = "refs.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
